---
title: "leaflet"
author: "Yi Li"
date: "3/11/2019"
output: html_document
---
# Learn How to Use Leaflet
Leaflet is a widely used open source JavaScript library used to build web mapping applications. First released in 2011, it supports most mobile and desktop platforms, supporting HTML5 and CSS3. 

## Install Leaflt from github
devtools: a package development tools for R.
```{r}
install.packages("devtools")
library(devtools)
devtools::install_github("rstudio/leaflet")
library(leaflet)
```

## Create a simple map

```{r}
m <- leaflet() %>%  # Create a map widget by calling leaflet
  addTiles() %>%    # Add default OpenStreetMap map tiles
  addMarkers(lat=40.7484405, lng=-73.9878531, popup="Empire State Building")
m  # Print the map widget to display it
```

## leaflet() widget = map + data
the leaflet() widget consists of two parts:
• map widget with initializing options to set up the map and map options to manipulate the attributes of the map
• data layer, including shape files and other associated data. You can add that right away or at a later time.

## Example 1: Police Complaints in NYC
https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Map-Year-to-Date-/2fra-mtpn
```{r}
# Get data on NYC police complaints
library(readr)
library(dplyr)
nypd <- read_csv("NYPD_Complaint_2018-06_Manhattan.csv")
# Remove values with missing coordinates
nypd <- nypd %>% filter(
  is.na(nypd$Latitude)==FALSE &
  is.na(nypd$Longitude)==FALSE)
# Plot
leaflet(nypd) %>% 
  addTiles() %>% # add the background map 
  addCircles(lng = ~Longitude, lat = ~Latitude) # lng = ~Longitude, lat = ~Latitude  can be removed
```
## To Change the background tiles...
Just copy the generated url from the JavaScript code into the addTiles layer or write the name of the provider into the addProviderTiles object
there are lots (~100) of other options at http://leaflet-extras.github.io/leaflet-providers/preview/
```{r}
(m = leaflet() %>% addTiles() %>% setView( lng = 2.34, lat = 48.85, zoom = 5 ) %>%
    addProviderTiles("NASAGIBS.ViirsEarthAtNight2012"))

```
## Web Map Service (WMS) tiles
Global population distribution (in 2015)
```{r}
leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 4) %>%  addWMSTiles(
"http://sedac.ciesin.columbia.edu/geoserver/wms",
layers = "gpw-v4:gpw-v4-population-count_2015",
options = WMSTileOptions(format = "image/png", transparent = TRUE),
tileOptions(tms = TRUE),
attribution = "SEDAC @ Columbia University")
```
##  Using providers available directly in leaflet
One way is to use the provider name
```{r}
leaflet(nypd) %>%
  addProviderTiles("Stamen.TonerLite") %>% # Using the Provider Name
  addCircles(col="orange")  
```
The other is to use the provider url
```{r}
leaflet(nypd) %>%
  addTiles('http://{s}.basemaps.cartocdn.com/dark_all/{z}/{x}/{y}.png') %>%
    # Using the provider URL (as in Javascript library)
  addCircles(col="orange")
```
## What is marker
It represent unique locations on the map.
## Try with nypd
```{r}
m <- leaflet(nypd) %>% 
     addTiles('http://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}.png') %>%
     setView(-73.9949344, 40.7179112, zoom = 14)
m %>% addCircles(col="orange")
```
### Data Wrangling
Select some offense indluding theft, harrassment, drugs etc.
```{r}
# Selected offenses
# sort(table(nypd$offense_desc1))
off <- nypd$offense_desc1 
offsel <- NA
offsel <- ifelse(off=="GRAND LARCENY" | off=="PETIT LARCENY" | off=="THEFT-FRAUD" | off=="THEFT OF SERVICES", "Theft", offsel)
offsel <- ifelse(off=="HARRASSMENT", "Harrassment", offsel)
offsel <- ifelse(off=="ASSAULT 3 & RELATED OFFENSES" | off=="FELONY ASSAULT", "Assault", offsel)
offsel <- ifelse(off=="DANGEROUS DRUGS", "Drugs", offsel)
offsel <- ifelse(is.na(offsel), "Other Crime", offsel)
# add a new column
nypd$offsel1 <- offsel
```
### Select a palette
```{r}
library(RColorBrewer)
pal = colorFactor("Set1", domain = nypd$offsel1) # Grab a palette
color_offsel1 = pal(nypd$offsel1)
m %>% addCircles(color = color_offsel1) # Set color of points
```

### Add legend and pop-pup
```{r}
# pop-pup with details
content <- paste("What:",nypd$offense_desc1,"<br/>",
                 "When:",nypd$date,"<br/>",
                 "Where:",nypd$premise_type,"<br/>",
                 "Severity:",nypd$crimelevel,"<br/>")

mpop2 <- m %>% addCircles(color=color_offsel1, popup = content) %>%
  addLegend(pal = pal, values = ~nypd$offsel1, title = "Crimes")

mpop2
```
## Cluster the points
To improve user experience, we can **cluster points** on the map. 
These **clusters map back into individual points** as users **zoom** into an area of the map.
I like this interactive plot. 
```{r}
mclust <- m %>% addCircleMarkers(color = color_offsel1, 
                       popup = content,
                       clusterOptions = markerClusterOptions()) #<<
mclust
```











